---
title: "Are we more civilized today?"
author: "Sara Marie"
date: "Last updated: `r Sys.time()`"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE)

```
#  Explore Homicide rates in NW Europe
In this exercise, you will load a Manuel Eisner's dataset to capture the developments in homicides from Middle Ages to 20th century. 

**Your task: Explore the data and create the visualisations as instructed in the script.** 

- **Submit to Github this script with solutions and its knitted version.**  
- **Submit to Brightspace a document with the (1) URL of this script and its knitted version in Github (2) answer of 250 words to the Final Question and (3) the Final Plot.**



This exercise is based on the dataset provided by [OurWorldInData project](https://ourworldindata.org/homicides) based at the Oxford University.

## The long-term trend in Homicides in Western Europe

Understanding how homicide rates have changed prior to the modern era requires the help of historians and archivists. Manuel Eisner, a criminology professor at the University of Cambridge, and his colleagues published the Historical Violence Database \href{https://cjrc.osu.edu/research/interdisciplinary/hvd}: a compilation of data on long-term trends in homicide rates, in addition to qualitative information such as the cause of death, perpetrator and victim. This database is limited to countries with relatively complete historical records on violence and crime – mainly Western Europe and the US.

Starting in the second half of the nineteenth century, these European regions have consistent police records of those accused of murder or manslaughter and annual counts of homicide victims. To go back further in time, reaching as far back as the thirteenth century, Eisner collected estimates (from historical records of coroner reports, court trials, and the police) of homicide rates made in over ninety publications by scholars.

Homicide rates – measured as the number of homicides per 100,000 individuals – up to 1990 are sourced from Eisner’s (2003) publication and the Historical Violence Database. 

Are homicide rates in Europe today lower or higher than in the past? Using the provided dataset, display and describe the long-run homicide rates for the five European regions: Italy, England, Germany, Netherlands and Scandinavia.

```{r library}
library(tidyverse)
```

## Load the available data from ourworldindata.org

You should always interrogate the source of your data. Who compiled it, from where, what is missing, how representative the data are? Check the data/Metadata.txt to learn about the data provenance.

```{r load-data}
Western_Europe <- read_csv("data/homicide-rates-across-western-europe.csv")
```

## Inspect the data

How clean and analysis-ready is the dataset? Do you understand what the column names represent? What is the difference between rate and homicide number?

```{r inspect}
head(Western_Europe)
```

Ok, the data look good except for the column `Homicide rate in Europe over long-term (per 100,000)` which is not very easy to work with.

- Use the `names()` function and assignment key to relabel this column to `homicides_per_100k` 

```{r relabel column}
# YOUR CODE
names(Western_Europe)[4] <- "homicides_per_100k"
names(Western_Europe)
```

Now, that you have looked at what the data looks like and what it represents, and streamlined it, let's see what big picture it contains.

## Let's see what the long-term trend is in homicides

- use `ggplot()` function and remember the`+` at the end of the line
- chose a meaningful `geom_......()`  for geometry (hint: points are not great)
- load `Year` on the `x` axis and `homicides_per_100k` column in y axis
- to color individual country entries consistently, assign the country column to the argument `color`.
- provide meaningful title and axis labels
- remember to change the `eval` flag so that the code chunk renders when knitted 

```{r plot-homicides, eval = TRUE}
# Visualization of homicides in Europe through time 
Western_Europe %>% 
  ggplot(aes(x = Year, y = homicides_per_100k, color = Entity)) +
  geom_line() +
  labs(
    title = "Homicide Rates in Western Europe Over Time",
    x = "Year",
    y = "Homicides per 100,000 People")+ 
  theme_bw()
```
Alright, the homicide rates should all be descending over time. What a comfort. But the viz is not super clear. Let's check the rates for individual countries.

## Uncouple the homicides of individual countries for easier view

You can visualize each country's trend separately by adding an extra argument to the ggplot, the `facet_wrap()` and feeding it the country column. If in doubt, check your ggplot tutorial and your country column name for exact usage.

- reuse the ggplot from the chunk above
- insert `facet_wrap()` after the specification of geometry to split countries in separate charts
- change the facet "layout" to two columns and three rows so that the trends are easier to see in horizontal layout. 


```{r eval = TRUE}
# Uncoupling the homicides of individual countries 
 Western_Europe %>% 
  ggplot(aes(x = Year, y = homicides_per_100k, color = Entity)) +
  geom_line() +
  facet_wrap(~ Entity, ncol = 2, nrow = 3) +
  labs(
    title = "Homicide Rates in Western Europe Over Time",
    x = "Year",
    y = "Homicides per 100,000 People")+
  theme_bw() 
```

## Compare the trends in homicide with the pattern of reign duration among Danish rulers through time.

- Load your Danish king dataset. Hopefully it is tidy and your years and duration of reign are all numeric.
- You need to have a consistent way of plotting the rulers' reign on the x axis, so I recommend you create a midyear column by calculating the middle of each monarch's rule (Hint: `midyear` = `endyear` - (`endyear`-`startyear`)/2)
- Start a ggplot plotting midyear on x axis and duration on y axis
- Try `geom_smooth()` for geometry 
- Provide meaningful labels and a title
- How would you characterize the trend compared to the homicides above?

```{r eval = TRUE}
# Loading tidyverse and ggplot2
library(tidyverse)
library(ggplot2)

# Loading my danish kings into R
danish_kings <- read.csv2("data/Danish Kings 2.csv")
glimpse(danish_kings)

#Changing values to numeric 
danish_kings <- danish_kings %>%
  mutate(
    Regerings.start = as.numeric(as.character(Regerings.start)),
    Regerings.slut = as.numeric(as.character(Regerings.slut))
  )

# calculation of midyear and duration
danish_kings <- danish_kings %>%
  mutate(
    midyear = Regerings.slut - (Regerings.slut - Regerings.start) / 2,
    duration = Regerings.slut - Regerings.start
  )

# Removing NA
danish_kings <- danish_kings %>% 
  drop_na(midyear, duration)

#Removing negative and 0-values 
danish_kings <- filter(danish_kings, duration > 0)

#adding the danish kings dataset  to ggplot 
ggplot(danish_kings, aes(x = midyear, y = duration)) +
  geom_point() +  
  geom_smooth(method = "loess", se = TRUE) +  
  labs(title = "Danish monarchs reign trough time",
       x = "Duration of reign (Years)",
       y = "Midyear of reign") +
  theme_minimal()

#How would you characterize the trend compared to the homicides above?
# - Man kan se at homicide raten er faldet en del over tid, samtidig med at danske monarkers regeringstid bliver længere og længere. Man kan diskutere om en mere stabil stat og monark har resulteret i et emre roligt og discplineret samfund, som derved har fået homicide raten til og falde.
 

```


**Final tasks:**

1) Plot: In the facetted plot above, move the legend from the current position on the side to below the facets, and label it "Country" instead of "Entity".

```{r eval = TRUE}
# Final task with name change 
 Western_Europe %>% 
  ggplot(aes(x = Year, y = homicides_per_100k, color = Entity)) +
  geom_line() +
  facet_wrap(~ Entity, ncol = 2, nrow = 3) +
  labs(
    title = "Homicide Rates in Western Europe Over Time",
    x = "Year",
    y = "Homicides per 100,000 People",
     color = "Country")+
  theme_bw() +
  theme(legend.position = "bottom")
```

```{r eval=TRUE}
# Final task with name change
Western_Europe %>% 
  ggplot(aes(x = Year, y = homicides_per_100k, color = Entity)) +
  geom_line() +
  labs(
    title = "Homicide Rates in Western Europe Over Time",
    x = "Year",
    y = "Homicides per 100,000 People",
    color = "Country"
  ) +
  theme_bw()+
  theme(legend.position = "bottom")
```



2) Rmarkdown: 

  - edit the author of the document, and convert 'Final Tasks' into heading #2 (like the other headings) 
  - add a `floating table of contents` to your Rmarkdown document, 
  - provide informative `chunk-names` and edit flags in your R chunks, and
  - automatically generate a `timestamp` to show when the document was last updated. (Hint: check the Rmarkdown episode in our Data Carpentry tutorial)

###  Final question: 

*In <250 words articulate your answer on the basis of the data visualisations to the following question*:** are we more civilized today?** *(Can you find an article or a book that discusses this topic?)*

- Som besvaret tidligere i opgaven, kan vi ud fra visualiseringerne se, at mord-raterne går hånd i hånd med kongehusets forlængede regeringstid. Der er en tydelig nedgang i mord-raterne, og man kan derfor sammenligne mord-raterne i Skandinavien med den voksende stabilitet i den danske kongehus. Som konsekvens af de danske monarkers længere regeringstid, kan vi altså vurdere ud fra statistikkerne, at man får et mere stabilt og roligt samfund. En mere stabil stat fører ofte til en disciplinering af samfundet, og hvis man går ud fra denne overvejelse, følger det fint visualiseringernes konklusioner.
Men en artikel af T. A. Ramesh kaldet “Are we really civilized” fra 2006, kommer med udlægningen om at ordet ”civilisering” stort set er en myte. Hans teori er, at civilisering er resultatet af en udvikling i menneskelig kultur. Han sammenligner oldtiden med det moderne samfund, ved at sige at civilisering altid har været en del af menneskelig kultur, men at den er kommet til udtryk på forskellige måder. Man kan altså ikke sige at mennesker i oldtiden ikke var civiliserede, idet arkæologien har vist at de havde ting som mønter, templer, havne, paladser, markedspladser, osv. Disse er alle tegn på et civiliseret samfund. 
I moderne tid kan vi nikke genkende til mange af disse ting. Vi som samfund i Europa, er bare begyndt at lægge mere vægt på værdier som frihed, rettigheder og opfindelser. Man kan ud fra denne artikel altså diskutere om vi er blevet mere “civiliserede” som samfund på baggrund af stat, eller om der bare er foregået en forandring i menneskelig kultur som har haft direkte indflydelse på menneskers opførsel indenfor de givne samfund.

