---
title: "Danish Kings week 10"
author: "Sara Marie"
date: "7. March 2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r library, echo = TRUE}
library(tidyverse)

read_csv("data/kings.csv")
```

The task here is to load your Danish Monarchs csv into R using the `tidyverse` toolkit, calculate and explore the kings' duration of reign with pipes `%>%` in `dplyr` and plot it over time.

## Load the kings

Make sure to first create an  `.Rproj` workspace with a `data/` folder where you place either your own dataset or the provided `kings.csv` dataset. 

1. Look at the dataset that are you loading and check what its columns are separated by? (hint: open it in plain text editor to see)

List what is the

separator:_______  

2. Create a `kings` object in R with the different functions below and inspect the different outputs.
-  `read.csv()`
-  `read_csv()`
-  `read.csv2()` 
-  `read_csv2()`
```{r eval=TRUE}
# FILL IN THE CODE BELOW and review the outputs
kings1 <- read.csv("data/kings.csv")

kings2 <- read_csv("data/kings.csv", na = "NULL")

kings3 <- read.csv2("data/kings.csv")

kings4 <- read_csv2( "data/kings.csv")
```
Answer: 
1. Which of these functions is a `tidyverse` function? Read data with it below into a `kings` object
- Den funktion vi kører videre med er kings2 fordi det er en tibble og derved også den som tidyverse funktionen fungere bedst i 

2. What is the result of running `class()` on the `kings` object created with a tidyverse function.
- når man kører class() på kings2 får man resultatet "spec_tbl_df" "tbl_df" "tbl" "data.frame" 
- Det betyder at kings2 er en tibble og ikke en dataframe sin kings 1

3. How many columns does the object have when created with these different functions?
- glimpse(kings2), på den måde ser vi at der er 11 kolonner

4. Show the dataset so that we can see how R interprets each column
- Vi kan se at R ser alle observationerne som "characters", altså kan R ikke se årstallene som tal. Hvilket kan skabe problemer 


```{r eval = TRUE}
# COMPLETE THE BLANKS BELOW WITH YOUR CODE, then turn the 'eval' flag in this chunk to TRUE.

kings <- kings2

class(kings)


head(kings)


glimpse(kings) 


```


## Calculate the duration of reign for all the kings in your table
You can calculate the duration of reign in years with `mutate` function by subtracting the equivalents of your `startReign` from `endReign` columns and writing the result to a new column called `duration`. But first you need to check a few things:

* Is your data messy? Fix it before re-importing to R
* Do your start and end of reign columns contain NAs? Choose the right strategy to deal with them: `na.omit()`, `na.rm=TRUE`, `!is.na()`



Create a new column called  `duration` in the kings dataset, utilizing the `mutate()` function from tidyverse. Check with your group to brainstorm the options.

```{r}
# Ved at bruge na.omit(kings) så fjerner den de NA's der skulle være i vores datasæt
na.omit(kings)

#en ny kolonne der hedder duration
kings_mid_year <- kings %>% 
  mutate(duration = End_year- Start_year) %>% 
  mutate(mid_year = Start_year + duration/2) %>% 
  glimpse()
```

## Calculate the average duration of reign for all rulers
Do you remember how to calculate an average on a vector object? If not, review the last two lessons and remember that a column is basically a vector. So you need to subset your `kings` dataset to the `duration` column. If you subset it as a vector you can calculate average on it with `mean()` base-R function. If you subset it as a tibble, you can calculate average on it with `summarize()` tidyverse function. Try both ways!

* You first need to know how to select the relevant `duration` column. What are your options?
* Is your selected `duration` column a tibble or a vector? The `mean()` function can only be run on a vector. The `summarize()` function works on a tibble.
* Are you getting an error that there are characters in your column? Coerce your data to numbers with `as.numeric()`.
* Remember to handle NAs: `mean(X, na.rm=TRUE)`


```{r}
# YOUR CODE
kings_mid_year %>% 
  ggplot(aes(x = mid_year, y = duration)) +
  geom_point()+
  geom_smooth()+
  theme_bw()
 
avg_duration <- mean(kings_mid_year$duration, na.rm=TRUE)

```

## How many and which kings enjoyed a longer-than-average duration of reign?

You have calculated the average duration above. Use it now to `filter()` the `duration` column in  `kings` dataset. Display the result and also  count the resulting rows with `count()`

```{r}
# YOUR CODE
long_reigning_kings <- kings_mid_year %>%
  filter(duration > avg_duration) %>%
  select(Name, duration) %>% 
  print(n = Inf)

```


## How many days did the three longest-ruling monarchs rule? 

* Sort kings by reign `duration` in the descending order. Select the three longest-ruling monarchs with the `slice()` function
* Use `mutate()` to create `Days` column where you calculate the total number of days they ruled
* BONUS: consider the transition year (with 366 days) in your calculation!

```{r}
# YOUR CODE
kings_mid_year%>% 
  arrange(desc(duration)) %>% 
  slice(1:3) %>% 
  mutate(Days = duration * 365 + floor(duration/4)) %>% 
  select(Name, duration, Days)

```


## Challenge: Plot the kings' duration of reign through time

What is the long-term trend in the duration of reign among Danish monarchs? How does it relate to the historical violence trends ? 

* Try to plot the duration of reign column in `ggplot` with `geom_point()` and  `geom_smooth()`
* In order to peg the duration (which is between 1-99) somewhere to the x axis with individual centuries, I recommend creating a new column `midyear` by adding to `startYear` the product of `endYear` minus the `startYear` divided by two (`startYear + (endYear-startYear)/2`).
* Now you can plot the kings dataset, plotting `midyear` along the x axis and `duration` along y axis
* BONUS: add a title, nice axis labels to the plot and make the theme B&W and font bigger to make it nice and legible!

```{r}
# YOUR CODE

```

And to submit this rmarkdown, knit it into html. But first, clean up the code chunks, adjust the date, rename the author and change the `eval=FALSE` flag to `eval=TRUE` so your script actually generates an output.
Well done!


