---
title: 'Text mining, sentiment analysis, and visualization by Sara Marie '
date: 'created on 22 November 2020 and updated `r format(Sys.time(), "%d %B, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = TRUE,
                      message = TRUE)

library(tidyverse)
library(here)

# For text mining:

library(pdftools)
library(tidytext)
library(textdata) 
library(ggwordcloud)
#Downloading Sentida
if(!require("devtools")) install.packages("devtools")

devtools::install_github("Guscode/Sentida")
library(Sentida)
force= TRUE
```

**Note:** for more text analysis, you can fork & work through Casey O’Hara and Jessica Couture’s eco-data-sci workshop (available here https://github.com/oharac/text_workshop)

### Downloading the Hunger games trilogy to RStudio
```{r get-document}

Hunger_games <- here("data/the_hunger_games_-_trilogy.pdf")
Hunger_games_text <- pdf_text(Hunger_games)
Hunger_games_text

```

Some things to notice:

- How cool to extract text out of a PDF! Do you think it will work with any PDF?
- Each row is a page of the PDF (i.e., this is a vector of strings, one for each page)
- The pdf_text() function only sees text that is "selectable"

Example: Just want to get text from a single page (e.g. Page 9)? 
```{r single-page}
Hunger_games_p112 <- Hunger_games_text[112]
Hunger_games_p112
```

See how that compares to the text in the PDF on Page 9. What has pdftools added and where?

From Jessica and Casey's text mining workshop: “pdf_text() returns a vector of strings, one for each page of the pdf. So we can mess with it in tidyverse style, let’s turn it into a dataframe, and keep track of the pages. Then we can use stringr::str_split() to break the pages up into individual lines. Each line of the pdf is concluded with a backslash-n, so split on this. We will also add a line number in addition to the page number."

### Some wrangling:

- Split up pages into separate lines (separated by `\n`) using `stringr::str_split()`
- Unnest into regular columns using `tidyr::unnest()`
- Remove leading/trailing white space with `stringr::str_trim()`

```{r split-lines}
Hunger_games_df <- data.frame(Hunger_games_text) %>% 
  mutate(text_full = str_split(Hunger_games_text, pattern = '\n')) %>% 
  unnest(text_full) %>% 
  mutate(text_full = str_trim(text_full)) 
Hunger_games_df

# More information: https://cran.r-project.org/web/packages/stringr/vignettes/regular-expressions.html
```

Now each line, on each page, is its own row, with extra starting & trailing spaces removed. 

### Get the tokens (individual words) in tidy format

Use `tidytext::unnest_tokens()` (which pulls from the `tokenizer`) package, to split columns into tokens. We are interested in *words*, so that's the token we'll use:

```{r tokenize}
Hunger_games_tokens <- Hunger_games_df %>% 
  unnest_tokens(word, text_full)
Hunger_games_tokens

# See how this differs from `Hunger_games_df`
# Each word has its own row!
```

Let's count the words!
```{r count-words}
Hunger_games_wc <- Hunger_games_tokens %>% 
  count(word) %>% 
  arrange(-n)
Hunger_games_wc
```

OK...so we notice that a whole bunch of things show up frequently that we might not be interested in ("a", "the", "and", etc.). These are called *stop words*. Let's remove them. 

### Remove stop words:

See `?stop_words` and `View(stop_words)`to look at documentation for stop words lexicons.

We will *remove* stop words using `tidyr::anti_join()`:
```{r stopwords}
Hunger_games_stop <- Hunger_games_tokens %>% 
  anti_join(stop_words) %>% 
  select(-Hunger_games_text)
Hunger_games_stop
```

Now check the counts again: 
```{r count-words2}
Hunger_games_swc <- Hunger_games_stop %>% 
  count(word) %>% 
  arrange(-n)
Hunger_games_swc
```

What if we want to get rid of all the numbers (non-text) in `Hunger_games_stop`?
```{r skip-numbers}
# This code will filter out numbers by asking:
# If you convert to as.numeric, is it NA (meaning those words)?
# If it IS NA (is.na), then keep it (so all words are kept)
# Anything that is converted to a number is removed

Hunger_games_no_numeric <- Hunger_games_stop %>% 
  filter(is.na(as.numeric(word)))
Hunger_games_no_numeric
```

### A word cloud of IPCC report words (non-numeric)

See more: https://cran.r-project.org/web/packages/ggwordcloud/vignettes/ggwordcloud.html

```{r wordcloud-prep}
# There are almost 2000 unique words 
length(unique(Hunger_games_no_numeric$word))

# We probably don't want to include them all in a word cloud. Let's filter to only include the top 100 most frequent?
Hunger_games_top100 <- Hunger_games_no_numeric %>% 
  count(word) %>% 
  arrange(-n) %>% 
  head(100)
Hunger_games_top100
```

```{r wordcloud}
# Installer stringer 
library(stringr)

# Fjern HTML-tags og andre ikke-alfanumeriske tegn
# FjernerHTML-tags
Hunger_games_top100$word <- str_remove_all(Hunger_games_top100$word, "<.*?>") 
#Fjerner ikke-alfanumeriske tegn
Hunger_games_top100$word <- str_replace_all(Hunger_games_top100$word, "[^[:alnum:]\\s]", "")  # 

Hunger_games_cloud <- ggplot(data = Hunger_games_top100, aes(label = word, size = n)) +
  geom_text_wordcloud() +
  theme_minimal()

Hunger_games_cloud
```



That's underwhelming. Let's customize it a bit:
```{r wordcloud-pro}
ggplot(data = Hunger_games_top100, aes(label = word, size = n)) +
  geom_text_wordcloud_area(aes(color = n), shape = "star") +
  scale_size_area(max_size = 12) +
  scale_color_gradientn(colors = c("purple","lightblue","hotpink")) +
  theme_minimal()
```

Cool! And you can facet wrap (for different reports, for example) and update other aesthetics. See more here: https://cran.r-project.org/web/packages/ggwordcloud/vignettes/ggwordcloud.html

### Sentiment analysis

First, check out the ‘sentiments’ lexicon. From Julia Silge and David Robinson (https://www.tidytextmining.com/sentiment.html):

“The three general-purpose lexicons are

  -  AFINN from Finn Årup Nielsen,
  -  bing from Bing Liu and collaborators, and
  -  nrc from Saif Mohammad and Peter Turney

All three of these lexicons are based on unigrams, i.e., single words. These lexicons contain many English words and the words are assigned scores for positive/negative sentiment, and also possibly emotions like joy, anger, sadness, and so forth. The AFINN lexicon assigns words with a score that runs between -5 and 5, with negative scores indicating negative sentiment and positive scores indicating positive sentiment.  The bing lexicon categorizes words in a binary fashion into positive and negative categories. The nrc lexicon categorizes words in a binary fashion (“yes”/“no”) into categories of positive, negative, anger, anticipation, disgust, fear, joy, sadness, surprise, and trust.  All of this information is tabulated in the sentiments dataset, and tidytext provides a function get_sentiments() to get specific sentiment lexicons without the columns that are not used in that lexicon."

Let's explore the sentiment lexicons. "bing" is included, other lexicons ("afinn", "nrc", "loughran") you'll be prompted to download.

**WARNING:** These collections include very offensive words. I urge you to not look at them in class.

"afinn": Words ranked from -5 (very negative) to +5 (very positive)
```{r afinn}
get_sentiments(lexicon = "afinn")
# You should be prompted to install lexicon - choose yes!

afinn_pos <- get_sentiments("afinn") %>% 
  filter(value %in% c(3,4,5))

afinn_pos
```

bing: binary, "positive" or "negative"
```{r bing}
get_sentiments(lexicon = "bing")
# You should be prompted to install lexicon - choose yes!

```

nrc:https://saifmohammad.com/WebPages/NRC-Emotion-Lexicon.htm
Includes bins for 8 emotions (anger, anticipation, disgust, fear, joy, sadness, surprise, trust) and positive / negative. 

**Citation for NRC lexicon**: Crowdsourcing a Word-Emotion Association Lexicon, Saif Mohammad and Peter Turney, Computational Intelligence, 29 (3), 436-465, 2013.

Now nrc:
```{r nrc}
get_sentiments(lexicon = "nrc")
# You should be prompted to install lexicon - choose yes!

```

Let's do sentiment analysis on the IPCC text data using afinn, and nrc. 


### Sentiment analysis with afinn: 

First, bind words in `ipcc_stop` to `afinn` lexicon:
```{r bind-afinn}
Hunger_games_afinn <- Hunger_games_stop %>% 
  inner_join(get_sentiments("afinn"))
Hunger_games_afinn
```

Let's find some counts (by sentiment ranking):
```{r count-afinn}
Hunger_games_afinn_hist <- Hunger_games_afinn %>% 
  count(value)

# Plot them: 
ggplot(data = Hunger_games_afinn_hist, aes(x = value, y = n)) +
  geom_col(aes(fill = value)) +
  theme_bw()
```

Investigate some of the words in a bit more depth:
```{r afinn-2}
# What are these '2' words?
Hunger_games_afinn2 <- Hunger_games_afinn %>% 
  filter(value == 2)
Hunger_games_afinn2
```

```{r afinn-2-more}
# Check the unique 2-score words:
unique(Hunger_games_afinn2$word)

# Count & plot them
Hunger_games_afinn2_n <- Hunger_games_afinn2 %>% 
  count(word, sort = TRUE) %>% 
  mutate(word = fct_reorder(factor(word), n))


ggplot(data = Hunger_games_afinn2_n, aes(x = word, y = n)) +
  geom_col() +
  coord_flip() +
  theme_bw()

#jeg ændre lidt på grafen så den bliver mere læselig og det giver et bedre overblik
Hunger_games_afinn2_n_top20 <- Hunger_games_afinn2_n %>% 
  top_n(20, n)

ggplot(data = Hunger_games_afinn2_n_top20, aes(x = word, y = n)) +
  geom_col() +
  coord_flip() +
  theme_bw() +
  theme(axis.text.y = element_text(size = 8))  #
```

Look back at the IPCC report, and search for "confidence." Is it typically associated with emotion, or something else? 

We learn something important from this example: Just using a sentiment lexicon to match words will not differentiate between different uses of the word...(ML can start figuring it out with context, but we won't do that here).

Or we can summarize sentiment for the report: 
```{r summarize-afinn}
Hunger_games_summary <- Hunger_games_afinn %>% 
  summarize(
    mean_score = mean(value),
    median_score = median(value)
  )
Hunger_games_summary 
```

The mean and median indicate *slightly* positive overall sentiments based on the AFINN lexicon. 

### NRC lexicon for sentiment analysis

We can use the NRC lexicon to start "binning" text by the feelings they're typically associated with. As above, we'll use inner_join() to combine the IPCC non-stopword text with the nrc lexicon: 

```{r bind-bing}
Hunger_games_nrc <- Hunger_games_stop %>% 
  inner_join(get_sentiments("nrc"))
Hunger_games_nrc
```

Wait, won't that exclude some of the words in our text? YES! We should check which are excluded using `anti_join()`:

```{r check-exclusions}
Hunger_games_exclude <- Hunger_games_stop %>% 
  anti_join(get_sentiments("nrc"))

# View(ipcc_exclude)

# Count to find the most excluded:
Hunger_games_exclude_n <- Hunger_games_exclude %>% 
  count(word, sort = TRUE)

head(Hunger_games_exclude_n)
```

**Lesson: always check which words are EXCLUDED in sentiment analysis using a pre-built lexicon! **

Now find some counts: 
```{r count-bing}
Hunger_games_nrc_n <- Hunger_games_nrc %>% 
  count(sentiment, sort = TRUE)


ggplot(data = Hunger_games_nrc_n, aes(x = sentiment, y = n)) +
  geom_col(aes(fill = sentiment))+
  theme_bw()
```

Or count by sentiment *and* word, then facet:
```{r count-nrc}
Hunger_games_nrc_n5 <- Hunger_games_nrc %>% 
  count(word,sentiment, sort = TRUE) %>% 
  group_by(sentiment) %>% 
  top_n(5) %>% 
  ungroup()

Hunger_games_nrc_gg <- ggplot(data = Hunger_games_nrc_n5, aes(x = reorder(word,n), y = n, fill = sentiment)) +
  geom_col(show.legend = TRUE) +
  facet_wrap(~sentiment, ncol = 2, scales = "free") +
  coord_flip() +
  theme_minimal() +
  labs(x = "Word", y = "count")

# Show it
Hunger_games_nrc_gg

# Save it
ggsave(plot = Hunger_games_nrc_gg, 
       here("figures","ipcc_nrc_sentiment.png"), 
       height = 8, 
       width = 5)

```

## Big picture takeaway

There are serious limitations of sentiment analysis using existing lexicons, and you should **think really hard** about your findings and if a lexicon makes sense for your study. Otherwise, word counts and exploration alone can be useful! 

## Your task

Taking this script as a point of departure, apply sentiment analysis on the Game of Thrones. You will find a pdf in the data folder. What are the most common meaningful words and what emotions do you expect will dominate this volume? Are there any terms that are similarly ambiguous to the 'confidence' above? 

### Credits: 
This tutorial is inspired by Allison Horst's Advanced Statistics and Data Analysis.
